A Study on Common Bug Fix Patterns in Rust

### 2022-10-29 sections/methodology.tex
minor edit

A code change is a modification in a program that changes the program's abstract syntax tree. A change pattern is a series of code changes that serves a specific purpose. This purpose can be improving code readability or maintainability (refactoring patterns), or rectifying program behaviour with respect to a specific functional or non-functional requirement (bug-fixing patterns). In this work, we are interested in latter group. Thus, from now on, by referring to code patterns we mean bug-fixing patterns.

The general idea for automatically finding the pervasive code patterns within a list of programs is to parse the programs and embed the important parts of the derived AST in fixed-size datapoints. Work in this vein makes the assumption that each cluster corresponds to one change pattern~\cite{hanam2016discovering,campos2019discovering,yang2022mining}.

Our pipeline for finding pervasive patterns in Rust programs also follows this idea. We decided to implement our pipeline completely in Python. To make this work, we need a way to get an AST as a Python dictionary. Syn\footnote{\url{https://crates.io/crates/syn}} is a Rust crate which aims to serve Rust procedural macros. However, it includes a parser which is suitable for our purposes; we simply had to write a preprocessor to transform Syn AST output into Python dictionaries.

To compute the contents of a change, we need two code revisions: the revision before the commit, and the revision after it. After parsing these two revisions, we will have two ASTs, encoded in two Python dictionaries. Tree diff algorithms can compute the differences between two arbitrary trees. If the trees are programs' abstract syntax trees, we call their diff an ASTDiff. We use dictdiffer to obtain ASTDiffs.

@@

Our target repositories were the top 18 most-starred open source Rust projects on GitHub. We mined (using Pydriller~\cite{spadini2018pydriller}) all the bug related commits and ran them through our pipeline. Then, we applied the DBSCAN clustering algorithm~\cite{ester1996density} on the resulting datapoints. In Section~\ref{sec:clustering_data} we justify our choice of DBSCAN for clustering, and how we tuned it to improve cluster quality.

>>>>

We define a code change to be a modification to a program's source code that changes the program's abstract syntax tree (AST). A change pattern is a set of code changes that serves a specific purpose. This purpose can be improving code readability or maintainability (refactoring patterns), or rectifying program behaviour with respect to a specific functional or non-functional requirement (bug-fixing patterns). In this work, we are interested in bug-fixing patterns. Thus, from now on, whenever we refer to code patterns, we mean bug-fixing patterns.

The general idea for automatically finding the pervasive code patterns within a list of programs is to parse the programs and, for us, to embed the important parts of the derived AST in fixed-size datapoints. Then, using a clustering algorithm, we can group these datapoints in different clusters. Work in this vein makes the assumption that each cluster corresponds to one class of change pattern~\cite{hanam2016discovering,campos2019discovering,yang2022mining}.

Our pipeline for finding classes of pervasive patterns in Rust programs also follows this idea. We decided to implement our pipeline completely in Python. To make this work, we need a way to get an AST as a Python dictionary. Syn\footnote{\url{https://crates.io/crates/syn}} is a Rust crate which aims to serve Rust procedural macros. However, it includes a parser which is suitable for our purposes; we simply had to write a preprocessor to transform Syn AST output into Python dictionaries.

To compute the contents of a change, we need two code revisions: the revision before the commit, and the revision after it. After parsing these two revisions, we will have two ASTs, encoded in two Python dictionaries. Tree diff algorithms can compute the differences between two arbitrary trees. When the trees are programs' abstract syntax trees, we call their diff an ASTDiff. We use dictdiffer to obtain ASTDiffs.

@@

Our target repositories were the top 18 most-starred open source Rust projects on GitHub. We mined (using Pydriller~\cite{spadini2018pydriller}) all the bug related commits and ran them through our pipeline. Then, we applied the DBSCAN clustering algorithm~\cite{ester1996density} on the resulting datapoints. In Section~\ref{sec:clustering_data} we justify our choice of DBSCAN for clustering, and we explain how we tuned it to improve cluster quality.
